---
title: "Calculation of growing season climate data"
author: "Elisabeth Vogel"
documentclass: article
output:
  html_document:
    df_print: paged
geometry: margin=1in
lang: english
fontsize: 11pt
classoption: landscape
---

### Calculation of growing season climate data

This collection of scripts and functions was developed to calculate climate conditions during the growing season (between planting and harvest) for four staple crops -- maize, rice, wheat and soy. This notebook provides an example of the calculation based on one crop, one sample crop calendar and a subset of the climate data.

### Data used

#### Crop calendars

The following crop calendars were used in this analysis:

  1) The Sacks et al. crop calendar
  2) The MIRCA2000 crop calendar
  3) The AgMIP v1.0 crop calendar

##### References:

1) Sacks et al.

- URL: https://nelson.wisc.edu/sage/data-and-models/crop-calendar-dataset/index.php
- Paper: Sacks, W.J., D. Deryng, J.A. Foley, and N. Ramankutty (2010). Crop planting dates: an analysis of global patterns. Global Ecology and Biogeography 19, 607-620. DOI: 10.1111/j.1466-8238.2010.00551.x.

2) MIRCA2000

- URL: https://www.uni-frankfurt.de/45218031/data_download
- Paper: Portmann, F. T., Siebert, S. & Döll, P. (2010): MIRCA2000 – Global monthly irrigated and rainfed crop areas around the year 2000: A new high-resolution data set for agricultural and hydrological modeling, Global Biogeochemical Cycles, 24, GB 1011, doi:10.1029/2008GB003435.

3) AgMIP harmonised crop calendar v1.0

- Paper: Elliott, J., C. Müller, D. Deryng, J. Chryssanthacopoulos, K.J. Boote, M. Büchner, I. Foster, M. Glotter, J. Heinke, T. Iizumi, R.C. Izaurralde, N.D. Mueller, D.K. Ray, C. Rosenzweig, A.C. Ruane, and J. Sheffield, 2015: The Global Gridded Crop Model Intercomparison: Data and modeling protocols for Phase 1 (v1.0). Geosci. Model Dev., 8, 261-277, doi:10.5194/gmd-8-261-2015.


#### Climate data

The following climate datasets are processed:

  1) CRU TS 3.23 dataset
  2) HADEX2 dataset 

##### References:

1) CRU TS 3.23

- URL: https://crudata.uea.ac.uk/cru/data/hrg/cru_ts_3.23/
- Paper: Harris, I., Jones, P.D., Osborn, T.J. and Lister, D.H. (2014), Updated high-resolution grids of monthly climatic observations – the CRU TS3.10 Dataset. Int. J. Climatol., 34: 623–642. doi: 10.1002/joc.3711

2) HADEX2

- URL: https://www.climdex.org/datasets.html
- Paper: Donat, M. G., L. V. Alexander, H. Yang, I. Durre, R. Vose, R. J. H. Dunn, K. M. Willett, E. Aguilar, M. Brunet, J. Caesar, B. Hewitson, C. Jack, A. M. G. Klein Tank, A. C. Kruger, J. Marengo, T. C. Peterson, M. Renom, C. Oria Rojas, M. Rusticucci, J. Salinger, A. S. Elrayah, S. S. Sekele, A. K. Srivastava, B. Trewin, C. Villarroel, L. A. Vincent, P. Zhai, X. Zhang and S. Kitching. 2013a. Updated analyses of temperature and precipitation extreme indices since the beginning of the twentieth century: The HadEX2 dataset, J. Geophys. Res. Atmos., 118, 2098–2118, http://dx.doi.org/10.1002/jgrd.50150.

### Calculation

#### Step 1: Prepare growing season calendars

```{r}
# Source function to prepare the crop calendars
source("code/01_prepare-growing-season-calendars_functions.R")

# Download crop calendars into: data/raw_data/crop_calendars
# For example, MIRCA 2000: https://www.uni-frankfurt.de/45218031/data_download

crop = "maize"
irrigation = "combined" # use the irrigation pattern with the largest area fraction per grid cell

# As example, prepare MIRCA et al. crop calendar
prepare_mirca2000_crop_calendar(crop = tolower(crop), irrigation = irrigation, verbose = TRUE)
```


##### Step 2: Prepare climate datasets

The climate data are re-arranged and re-gridded to be used with the crop calendars. The bash script requires NCO and CDO for the processing of netcdf files. Further information on how to install NCO and CDO can be found under:

- NCO: http://nco.sourceforge.net/
- CDO: https://code.mpimet.mpg.de/projects/cdo

```{bash}
bash code/02_prepare-climate-input-data.sh
```


##### Step 3: Calculate growing season climate data


```{r}
# Source functions to calculate the growing season climate statistics
source("code/03_calculate-growing-season-climate_functions.R")

###############################################################################
# Example: Prepare growing season data for temperature in the CRU TS 3.23 dataset

crop_calendar_path = "data/processed_data/crop_calendars/mirca2000_crop_calendar"
cru_ts_path = "data/processed_data/climate_data/cru_ts_323"
out_path = "data/processed_data/growing_season_climate/cru_ts_323"
dir.create(out_path, showWarnings = FALSE, recursive = TRUE)

crop = "maize"
var = "tmp"
unit = "degrees Celsius"
longname = "temperature"
irrigation = "combined"

crop_calendar_nc = file.path(crop_calendar_path, sprintf("mirca_crop_calendar_%s_%s.nc", crop, irrigation))
climate_nc = file.path(cru_ts_path, sprintf("cru_ts323_%s_1961_2014.nc", var))
output_nc = file.path(out_path, sprintf("%s_%s_cru_ts_323_%s_gs.nc", crop, irrigation, var))

gs = calculate_growing_season_climate(crop_calendar_nc, climate_nc, output_nc, var, unit, longname,
                                 time_agg = "month", verbose = TRUE)
```
